Untwisting the Tetrahelix (v0.4)

### 2017-02-15 doc/TetrahelixUntwist.tex
proof simplifcations

From Robert Gray's site, repeating formula by H.S.M. Coxeter:
\[
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos(n \theta $\\
   $ r_{bc} \sin(n \theta) $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\]
where $n$ represents each integer numbered node in succession.

@@

@@

 \subsection{Recent Notes}

 Any minimax length optimal tetrahelix must have the same pitch on each helix, the length of one set of members grows unboundedly with
 each revolution. So in seeking a minimax optimal tetrahelix we only consider those with the same torsion.

@@

 \subsection{Very bad attempts to prove things}

 First prove that any structure isomorphic to the BC-Helix of unbounded extent in which each rail lies on a helix has certain
 properties if it it member length minimax optimal.

 The pitches must be equal, or some member grows without bound.

 If the axes are not parallel, some member grows without bound.

 We can therefore think in terms of O,G, P displacement.

 We need to prove that the three axes are coincident.

 Suppose there is one helix that is not conicident with the other two.  So long as $\rho \neq 0$, there is some value of $\kappa$
 makes the lenth to any node longer and shorter. Then the distance between nodes depends on $\kappa$, unless $\rho = 0$ or $\rho = \pi$.
 One of these makes the longest distance longer and the shortest distance shorter, so a superior solution occurs when the a
 axes conincide.  This is actually correct but poorly written.

  Note: This is a new, simpler proof attempt.
  \begin{theorem}
    Suppose that all edge vertices are on one of three helices of the same positive pitch and the same radius with the
    same shared axes about the $z$ axis differing by $120\degree$ about $z$.
    However, we do not assume that the three helices are equidistant.
    Without loss of generality, consider $A$ and $B$ to be to be on the same axis, $C$ to be on a second axis, and $D$ to be on the final.
    Without loss of generality assume: $A.z \leq C.z \leq D.z \leq B.z$.
    Then the minimum maximum difference between any two nodes is acheived when
    $ \overline{\rm AD} =  \overline{\rm BC} $ and $ \overline{\rm AC} =  \overline{\rm CD}, \overline{\rm DB} $.
  \end{theorem}
Note: Change this to PQR. Also, may be better to use $\alpha$ instead of $\rho$ in this proof.
  \begin{proof}
\begin{align*}
  O  &= C.z - A.z \\
  G &= D.z - C.z \\
  P &=  B.z - D.z
\end{align*}

Using the cartesian distance formula:
\begin{align*}
  \overline{\rm CD}  &= \sqrt{G^2 + (r\sin{0 \rho} -  r\sin{G \rho})^2 + (r\cos{0 \rho} - r\cos{G \rho})^2} \\
    \overline{\rm CD}  &= \sqrt{G^2 + (0 -  r\sin{G \rho})^2 + (r - r\cos{G \rho})^2} \\
    \overline{\rm CD}  &= \sqrt{G^2 + r^2(\sin{G \rho})^2 + r^2(1 - \cos{G \rho})^2} \\
    \overline{\rm CD}  &= \sqrt{G^2 + r^2(1 + \sin^2{G \rho}  - \cos^2{G \rho})} \\    
\end{align*}

Note: According to Wolfram Alpha, $\sin^2{G \rho}  - \cos^2{G \rho} = -\cos{2 G \rho}$

So, for all O,G,P,
\begin{align*}
  \overline{\rm AB}  &= 1 \\        
  \overline{\rm CD}  &= \sqrt{G^2 + r^2(1 + \sin^2{G \rho}  - \cos^2{G \rho})} \\
  \overline{\rm AC}  &= \sqrt{O^2 + r^2(1 + \sin^2{O \rho}  - \cos^2{O \rho})} \\
  \overline{\rm BD}  &= \sqrt{P^2 + r^2(1 + \sin^2{P \rho}  - \cos^2{P \rho})} \\
  \overline{\rm AD}  &= \sqrt{(O+G)^2 + r^2(1 + \sin^2{(O +G) \rho}  - \cos^2{(O+G) \rho})} \\
  \overline{\rm BC}  &= \sqrt{(P+G)^2 + r^2(1 + \sin^2{(P +G) \rho}  - \cos^2{(P+G) \rho})} \\      
\end{align*}

We show that the maximum difference between these variables for a given $r$ is minimized when $O=G=P=1/3$.
First we show any configuration in which $O \neq P$ implies a solution at least as good in terms of minimax difference in which $O = P$.
Let 
\begin{align*}
  \frac{O' + P'}{2} &= \frac{1}{2} + \delta
\end{align*}
Then any configuration $O' - \delta/2, P' - \delta/2$ has a minimax at least as good as $O',P'$, because, if $\delta > 0$, then
$\overline{\rm A'C'}$ is the longest segment of the four $\overline{\rm A'C'}, \overline{\rm A'D'}, \overline{\rm B'C'} =  \overline{\rm B'D'} $.
$\overline{\rm B'D'}$ is the shortest of this set.
\begin{align*}
  \overline{\rm A'D'} &> \overline{\rm B'C'}  \\
   \overline{\rm B'D'} &< \overline{\rm A'C'}  \\  
   \overline{\rm AB} &=   \overline{\rm A'B'}   \\
   \overline{\rm CD} &=   \overline{\rm C'D'}   \\
  \\
    \overline{\rm AC}&=   \overline{\rm BD}   \\
   \overline{\rm AD} &=   \overline{\rm BC}   \\
  \\  
   \overline{\rm AC} &<  \overline{\rm A'C'} \\
   \overline{\rm AD} &< \overline{\rm A'D'}   \\
   \overline{\rm BC} &>  \overline{\rm B'C'} \\
   \overline{\rm BD} &> \overline{\rm B'D'}   \\
\end{align*}
So the maximum $\overline{\rm AD}$ is at most the old maximum. The minimum $\overline{\rm AC}$ is at least the old minimum of our set of four.
If the maximum or minimum of a subset is not worsened, the maximum or minimum of the total set cannot be worsened. Therefore $O' - \delta/2, P' - \delta/2$
is no worse than $O', P'$. So any mimimax solution is no better than one in which $(O + P)/2 = 1/2$, or $O = P$.

  Any tetrahelix $ABCD$ of edge lengths $AD = u, AB = BC = CD = v , AC = BD = w$, where $1 \leq u \leq v$ is mimimax optimal.
  Any one tetrahedron in a tetrahelix has $1$ rail edge, $2$ mid-length edges connected to the rail and $2$ long edges connected to the rail.
  The edge opposite of the rail edge is a mid-length edge.
    \end{proof}

\begin{theorem}
  \label{tetrahelixoptimality}
  Any one tetrahedron in a minimax-optimal tetrahelix $ABCD$ of edge lengths $\overline{\rm AD} = 1, \overline{\rm AB} = \overline{\rm BC} = \overline{\rm CD} = u , \overline{\rm AC} = \overline{\rm BD} = v$, where $1 \leq u \leq v$.
  Any one tetrahedron in a tetrahelix has $1$ rail edge, $2$ mid-length edges connected to the rail and $2$ long edges connected to the rail.
  The edge opposite of the rail edge is a mid-length edge.
\end{theorem}

\begin{proof}
The edge opposite of the rail edge is a mid-length edge because between any two rails the edges alteranate mid-length and long.
\end{proof}

\begin{corollary}
  \label{eventhirds}
  Any minimax optimal tetrahelix has its colored edges spaced at 1/3 the height of
  a tetrahedra along the z-axis, where the height is the z-distance of the rail edge.
  \end{corollary}

\begin{proof}
  Let the variables $O,G,P$ represent the z-axis distance from the
  $R_0$ to $Y_0$, $Y_0$ to $B_0$, and
  $B_0$ to $R_0$ respectively. Then $O+G+P = h$. But $O=G=P$, since each edge between
  their nodes spans two rails of the same distance with an edge of the same length,
  so that the $z$-distance between each node is the same. So $O=G=P=h/3$.
\end{proof}

Note that based on Theorems \ref{tetrahelixoptimality} and \ref{eventhirds}, we are justified in classifying edge lengths as \emph{rail},\emph{one-hop}, or
\emph{two-hops}. The one-hop edges are the edges between closest on the $z$-axis, and the two-hop edges are those that hop over a vertex.

Every optimal tetrahelix has vertices lying on helices expressible in the form:
\[
V_{optimal}(n,c) =
\left [
  \begin{tabular}{c}
   $ r \cos(n \alpha +  c 2 \pi /3)$\\
   $ r \sin(n \alpha +  c 2 \pi /3) $\\
   $ \frac{d(n +c / 3)}{3}   $
  \end{tabular}
\right ]
\]
where we have not yet investigated in the general case the relationships beteween $\alpha$, $r$, and $d$ in this formulation.
However, we understand that when $\alpha = 0$, the helices are degenerate, having curvature of $0$, and
we have the equitetrabeam.

@@

By Theoerm \ref{eventhirds}, we can compute the (at most) three edge-lengths of an optimal
tetrahelix by (where $dist$ is the cartesian distance function):
\begin{align*}
  \text{rail} &= dist(H_{general}(n,c,\rho,h_{\rho}),H_{general}(n+1,c,\rho,d_{\rho})) = 1 \\
  \text{one-hop} &= dist(H_{general}(n,c,\rho,h_{\rho}),H_{general}(n,c+1,\rho,d_{\rho}))  \\
  \text{two-hops} &= dist(H_{general}(n,c,\rho,h_{\rho}),H_{general}(n,c+2,\rho,d_{\rho}))  \\  
\end{align*}
Where are invarinat for all $n$ and $c$.

@@

>>>>

From Robert Gray's site, repeating formula by H.S.M. Coxeter:
\[
V(n) =
\left [
  \begin{tabular}{c}
   $ r_{bc} \cos(n \theta) $\\
   $ r_{bc} \sin(n \theta) $\\
   $ n h_{bc}  $
  \end{tabular}
  \right ],
\text{where:}
  \begin{tabular}{c}
 $ r_{bc} = \frac{3\sqrt{3}}{10} $\\
 $ h_{bc} = 1/\sqrt{10} $ \\
 $ \theta = \arccos(-2/3) $ \\
  \end{tabular}      
\]
where $n$ represents each integer numbered node in succession.

@@

\section{Optimal Tetrahelices Have Evenly Spaced Vertices}

We use the term \emph{tetrahelix} to mean any structure made of vertices and edges which is isomorphic to the BC helix,
in which the vertices lie on three helices, 
no matter what lengths the edges take on. One could consider various definitions of optimality for a tetrahelix,
but the must useful to an enginner is to minimize the maximum difference between any two edges. We call a
tetrahelix \emph{member-length minimax optimal} if there is not tetrahelix of the same radius and pitch with a smaller
maximum edge length difference.

Consider a tetrahelix isomporphic to an BC helix of unbounded extent. If all three rails do not have same pitch, there
is an edge of unbounded length. So we are justified in talking about the pitch of the tetrahelix, even though we think
of it as three helices. Similarly, if the axes are not parallel, there is an edge of unbounded length in the structure,
so it is not optimal.

Consider any tetrahelix in which the axes of the helices are parallel to $z$ but not coincident, but in which all three helices have the same
radius. Consider the projection along the $z$ axis, which will be a figure of dots and connecting segments in the $xy$ plane. The convex
hull for any one helix projection will be a circle (if its pitch is irrational) or an polygon if rational. So long as its rail angle is
not $0$, it will have at least two points in the plane. One of these will be longer and will be made shorter by moving that helix closer
to the midpoint of the other two in the $xy$ plane. Even if the rail angle is $\pi$, there will be two vertices for each rail, each of
which is connected to both vertices of the other two rails. At lesast one of these lines is a longest line which will be made shorter by moving
that helix in the $xy$ plane closer to the midpoint of the other two.

So any any optimal tetrahelix with a rail-angle of greater than $0$, that is, with any curvature, will have conincident axes.

We call a tetrahelix with $0$ curvature a \emph{tetrabeam}, and consider it as an important special case and an endpoint of a continuum
of tetrahelices of decreasing curvature.

@@

@@

 \begin{theorem}
   \label{eventhirds}
   A tetrahelix of a given radius and height $h$ in which all nodes are evenly spaced at $h/3$ intervals on the $z$ axis is optimal.
     Any one tetrahedron in a tetrahelix has $1$ rail edge, $2$ one-hop edges connected to the rail and $2$ two-hop edges connected to the rail.
  The edge opposite of the rail edge is a one-hop edge.
  \end{theorem}

   In principle in any tetrahelix where the three helices have any displacement along the $z$ axis there are 9 distinct edge classes:
   3 rail edges, a one-hop length between eachof 3 rails, and two hop length between each of three rails, where the two-hop length is a maximum
   length between rails (which could equal the one-hop length.) However, we have already shown the pitches and the rail lengths are equal
   in any optimal tetrahelix.

    Consider the tetrahelix in which the vertices  are evenly spaced at $h/3$ intervals on the $z$ axis. Every edge is either a rail edge, or it makes one
    hop, or it makes two hops. All of the one-hop edges are equal length.  All of the two-hop edges are equal length.

    Any displacemnt along the $z$ axis of any rail increases the length of one or two two-hop edges and shortens the length of one or two one-hop edges.
    This increases the minimax distance no matter what the rail edge length is, since all rail edges are the same length. Therefore, an evenly spaced
    tetrahelix is the unique optimal for any given radius and height.
  \end{proof}

Note that based on \ref{eventhirds}, we are justified in classifying edge lengths as \emph{rail},\emph{one-hop}, or
\emph{two-hops}. The one-hop edges are the edges between closest on the $z$-axis, and the two-hop edges are those that hop over a vertex.

By \ref{eventhirds} every optimal tetrahelix has vertices lying on helices expressible in the form:
\[
V_{optimal}(n,c) =
\left [
  \begin{tabular}{c}
   $ r \cos(n \alpha +  c 2 \pi /3)$\\
   $ r \sin(n \alpha +  c 2 \pi /3) $\\
   $ \frac{d(n +c / 3)}{3}   $
  \end{tabular}
  \right ],
\text{where:}
\begin{tabular}{c}
  $c \in \{0,1,2\}$
  \end{tabular}      
\]
where we have not yet investigated in the general case the relationships beteween $\alpha$, $r$, and $d$ in this formulation.
However, we understand that when $\alpha = 0$, the helices are degenerate, having curvature of $0$, and
we have the equitetrabeam.

@@

By Theoerm \ref{eventhirds}, we can compute the (at most) three edge-lengths of an optimal
tetrahelix by (where $dist$ is the cartesian distance function):
\begin{align*}
  \text{rail} &= dist(H_{general}(n,c,\rho,h_{\rho}),H_{general}(n+1,c,\rho,d_{\rho})) = 1 \\
  \text{one-hop} &= dist(H_{general}(n,c,\rho,h_{\rho}),H_{general}(n,c+1,\rho,d_{\rho}))  \\
  \text{two-hops} &= dist(H_{general}(n,c,\rho,h_{\rho}),H_{general}(n,c+2,\rho,d_{\rho}))  \\  
\end{align*}
Which are invarinat for all $n$ and $c$.

\begin{align*}
  \text{one-hop} &= dist(H_{general}(n,c,\rho,h_{\rho}),H_{general}(n,c+1,\rho,d_{\rho}))  \\
  \text{one-hop} &= dist(H_{general}(0,0,\rho,h_{\rho}),H_{general}(0,1,\rho,d_{\rho}))  \\  
  \text{one-hop}  &= \sqrt{d_{\rho}^2/9 + (r\sin{\frac{\rho(3\cdot 0 + 0)}{2\pi}} - r\sin{\frac{\rho(3\cdot 0 + 0+1)}{2\pi}})^2  +
    (r\cos{\frac{\rho(3\cdot0+0)}{2\pi}} - r\cos{\frac{\rho(3\cdot + 0+1)}{2\pi}})^2
  \text{one-hop}  &= \sqrt{d_{\rho}^2/9 + (r\sin{\frac{\rho(0)}{2\pi}} - r\sin{\frac{\rho(1)}{2\pi}})^2  +
    (r\cos{\frac{\rho(0)}{2\pi}} - r\cos{\frac{\rho(1)}{2\pi}})^2
  } \\  
\end{align*}

@@

    \section{Old Proofs}

We can create the distance between representative nodes as:
\begin{align*}
  \overline{\rm AB}  &= 1 \\        
  \overline{\rm CD}  &= \sqrt{G^2 + r^2(1 + -\cos{2 G \rho})} \\
  \overline{\rm AC}  &= \sqrt{O^2 + r^2(1 + \sin^2{O \rho}  - \cos^2{O \rho})} \\  
  \overline{\rm AC}  &= \sqrt{O^2 + r^2(1 + -\cos{2 O  \rho})} \\
  \overline{\rm BD}  &= \sqrt{P^2 + r^2(1 + -\cos{2 P \rho}))} \\
  \overline{\rm AD}  &= \sqrt{(O+G)^2 + r^2(1 + -\cos{2 (O + G)  \rho})} \\
  \overline{\rm BC}  &= \sqrt{(P+G)^2 + r^2(1 + -\cos{2 (P + G) \rho})} \\      
\end{align*}

Using the cartesian distance formula:
\begin{align*}
  \overline{\rm CD}  &= \sqrt{G^2 + (r\sin{0 \rho} -  r\sin{G \rho})^2 + (r\cos{0 \rho} - r\cos{G \rho})^2} \\
    \overline{\rm CD}  &= \sqrt{G^2 + (0 -  r\sin{G \rho})^2 + (r - r\cos{G \rho})^2} \\
    \overline{\rm CD}  &= \sqrt{G^2 + r^2(\sin{G \rho})^2 + r^2(1 - \cos{G \rho})^2} \\
    \overline{\rm CD}  &= \sqrt{G^2 + r^2(1 + \sin^2{G \rho}  - \cos^2{G \rho})} \\
    \overline{\rm CD}  &= \sqrt{G^2 + r^2(1 + -\cos{2 G \rho})} \\        
\end{align*}
Note: $\sin^2{x \rho}  - \cos^2{x \rho} = -\cos{2 x \rho}$

So, for all O,G,P,
\begin{align*}
  \overline{\rm AB}  &= 1 \\        
  \overline{\rm CD}  &= \sqrt{G^2 + r^2(1 + -\cos{2 G \rho})} \\
  \overline{\rm AC}  &= \sqrt{O^2 + r^2(1 + \sin^2{O \rho}  - \cos^2{O \rho})} \\  
  \overline{\rm AC}  &= \sqrt{O^2 + r^2(1 + -\cos{2 O  \rho})} \\
  \overline{\rm BD}  &= \sqrt{P^2 + r^2(1 + -\cos{2 P \rho}))} \\
  \overline{\rm AD}  &= \sqrt{(O+G)^2 + r^2(1 + -\cos{2 (O + G)  \rho})} \\
  \overline{\rm BC}  &= \sqrt{(P+G)^2 + r^2(1 + -\cos{2 (P + G) \rho})} \\      
\end{align*}

Then we can simplify our distance equations:
\begin{align*}
  \overline{\rm AB}  &= 1 \\        
  \overline{\rm CD}  &= \sqrt{G^2 + r^2(1 + -\cos{2 G \rho})} \\
  \overline{\rm AC} &=   \overline{\rm BD} &= \sqrt{P^2 + r^2(1 + -\cos{2 P \rho}))} \\
  \overline{\rm AC} &=   \overline{\rm BD} &= \sqrt{((1-G)/2)^2 + r^2(1 + -\cos{ \rho (1-G) }))} \\  
  \overline{\rm AD} &=  \overline{\rm BC} &=  \sqrt{((P+G)^2 + r^2(1 + -\cos{2 (P + G) \rho})} \\
  \overline{\rm AD} &=  \overline{\rm BC} &=  \sqrt{((1+ G)/2 )^2 + r^2(1 + -\cos{\rho (1 + G)})} \\        
\end{align*}

When $G = 0$, $\overline{\rm CD} = 0$. Our solution is improved as we increase $G$ until either $ \overline{\rm CD} = \overline{\rm AC}$, or the
quantity $ \overline{\rm BC} - \overline{\rm CD} $ starts increasing rather than decreasing (that is, when the derivative is positive.)
